---
title: "descriptives"
author: "Adelaida Barrera"
date: "3/7/2021"
output: html_document
---

```{r, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(summarytools)
library(kableExtra)
load("data/clean_data.Rdata")

```


```{r}
variables <- c("Experienced control by male partner (percentage)", 
               "Experienced emotional violence by male partner (percentage)", 
               "Experienced less severe phyiscal violence by male partner (percentage)",
               "Experienced severe physical violence by male partner (percentage)",
               "Years of education (women)", 
               "Years of education (men)",
               "Female has higher education level than male (percentage)",
               "Female has more years of eucation than male (percentage)",
               "Difference in years of education betwen female and male",
               "Female earns more money than male (percentage)", 
               "Age (women)", 
               "Age (men)",
               "Number of children born (women)",
               "Wealth index: 'poorest' or 'poor (percentage)",
               "Wealth index: 'middle' (percentage)",
               "Wealth index: 'rich' or 'richest (percentage)",
               "Afro-colombians (percentage)", 
               "Indigenous people (percentage)",
               "Other minorities (percentage)", 
               "Non-minorities (percentage)", 
               "Urban households (percentage)", 
               "Rural households (percentage)")

# Function to crete rows of descriptives table 

Create_row <- function(x, data){data %>% 
    summarise(mean = round(mean({{x}}), 2), 
              sd = round(sd({{x}}), 2),
              min = min({{x}}), 
              max = max({{x}}),
              range = paste0(min, "-" , max)) %>% 
    select(mean, sd, range) %>% rbind(descr_tbl, .)
}

mean <- c()
sd <- c()
range <- c()

descr_tbl <- tibble(mean, sd, range)

descr_tbl <- Create_row(x = ever_control, data)
descr_tbl <- Create_row(x = ever_emotional, data)
descr_tbl <- Create_row(x = ever_lesssevere, data)
descr_tbl <- Create_row(x = ever_severe, data)

descr_tbl <- Create_row(x = eduyear_f, data)
descr_tbl <- Create_row(x = eduyear_m, data)

descr_tbl <- Create_row(x = rel_level_binary, data)
descr_tbl <- Create_row(x = rel_years_binary, data)
descr_tbl <- Create_row(x = diff_years, data)

data_trim <- data %>% select(rel_income) %>% drop_na() 
descr_tbl <- Create_row(x = rel_income, data_trim)

descr_tbl <- Create_row(x = v012, data) # female respondents' age 

descr_tbl <- Create_row(x = mv012, data) # male respondents' age

descr_tbl <- Create_row(x = v201, data) # children ever born 

descr_tbl <- Create_row(x = poor_dummy, data) # wealth index:
descr_tbl <- Create_row(x = middle_dummy, data)
descr_tbl <- Create_row(x = rich_dummy, data)

descr_tbl <- Create_row(x = afro_dummy, data) 
descr_tbl <- Create_row(x = indigenous_dummy, data) 
descr_tbl <- Create_row(x = minorities_dummy, data) 
descr_tbl <- Create_row(x = mestizo_dummy, data) 

descr_tbl <- Create_row(x = urban_dummy, data) 
descr_tbl <- Create_row(x = rural_dummy, data) 

descr_tbl <- descr_tbl %>% mutate(Variable = variables) %>% relocate(Variable)

descr_tbl <- descr_tbl %>% mutate()

descr_tbl_prnt <- descr_tbl %>% 
  kbl(caption = "Desrciptive statistics for all relevant variables",
      format = "latex") %>% 
  kable_classic_2(latex_options =c("scale_down")) %>% 
  pack_rows("Violence", 1,4) %>% 
  pack_rows("Education", 5, 8) %>% 
  pack_rows("Relative empowerment", 9, 12) %>% 
  pack_rows("Other", 13, 15) 


kable_as_image(descr_tbl_prnt, filename =  "Table.2", file_format = "png")

descr_tbl

rm(mean, sd, range, variables, )

```
```{r}


level_plot <- data %>% 
  select(v002, factor_edulev_f, factor_edulev_m)  %>%
  pivot_longer(cols = c(factor_edulev_f, factor_edulev_m),
               names_to = "gender", 
               values_to = "edu_level") %>% 
  group_by(gender, edu_level) %>% 
  summarize(frequency = n()) %>% 
  mutate(`%` = (round(frequency/sum(frequency), 2)),
         gender = if_else(gender == "factor_edulev_f", "Women", "Men")) %>% 
  #   edu_level = case_when(edu_level == "1-less than primary" ~ "Less than primary",
  #                        edu_level == "2-primary" ~ "Primary",
  #                       edu_level == "3-secondary" ~ "Secondary",
  #                      edu_level == "4-higher" ~ "Higher education")) %>% 
  ggplot(aes(x = gender,
             y = `%`,
             fill = edu_level, 
             label = `%`)) +
  geom_col(position = "dodge") + 
  geom_text(position = position_dodge(width = 0.9),  vjust = -1,
            size = 4) +
  theme_minimal()  +
  scale_fill_ordinal() +
  scale_y_continuous(limits = c(0, 0.5)) +
  labs(title = "Figure 1. Educational attainment distribution by gender \n",
       x = "",
       y = "%",
       fill = "Education Level") +
  theme(legend.position = "bottom")


level_plot 
```

```{r edu_diff_plot, fig.height=5, fig.width=7}

data %>% 
  summarize(edu_average = ((eduyear_f + eduyear_m)/2),
            diff_edu = eduyear_m - eduyear_f,
            factor_diff_edu = 
              case_when(diff_edu > 1 ~ "3- Man is more educated",
                        diff_edu == 0 ~ "2- Education is the same",
                        diff_edu < 1 ~ "1- Woman is more educated")) %>% 
  na.omit() %>% 
  ggplot(aes(x = diff_edu,
             fill = factor_diff_edu)) + 
  geom_histogram(binwidth = 1.8) +
  theme_minimal() +
  labs(title = "Figure 2. Distribution of education gap in couples",
       y = "Count",
       x = "Education difference in years (Men - Women)",
       fill = "") +
  scale_fill_ordinal(begin = 0.1, end = 0.8) + 
    theme(legend.position = "bottom")


```

```{r age_plot, fig.height=4, fig.width=7}

age_plot <- data %>% 
  select(v002, age_f, age_m)  %>%
  pivot_longer(cols = c(age_f, age_m),
               names_to = "gender", 
               values_to = "age") %>% 
  mutate(gender = if_else(gender == "age_f", "Women", "Men")) %>% 
  ggplot(aes(x = age,
             fill = gender)) +
  geom_density(alpha = 0.6, size = 0) + 
  theme_minimal()  +
  scale_fill_ordinal() +
  labs(title = "Age distribution by gender \n",
       x = "Age (years)",
       y = "Density",
       fill = "Gender") + 
    theme(legend.position = "bottom") 
  

age_plot

```

```{r age_diff_plot, fig.height=4, fig.width=7,}

data %>% 
  summarize(age_average = ((age_m + age_f)/2),
            diff_age = age_m - age_f,
            factor_diff_age = 
              case_when(diff_age > 1 ~ "3-Man is older",
                        diff_age == 0 ~ "2-Same age",
                        diff_age < 1 ~ "1-Woman is older")) %>% 
  na.omit() %>% 
  ggplot(aes(x = diff_age,
             fill = factor_diff_age)) + 
  geom_histogram(binwidth = 1.9) +
  theme_minimal() +
  labs(title = "Figure 3. Distribution of age gap within couples",
       y = "Count",
       x = "Age difference in years (Man - Woman)",
       fill = "") +
  scale_fill_ordinal(direction = 1, begin = 0.1, end = 0.8) +
  scale_y_continuous(limits = c(0,2000)) + 
    theme(legend.position = "bottom")

```

```{r wealth_bw_plot, fig.height = 5, fig.width = 8}


# Wealth and female bread winners

data %>% 
  select(v002, wealth_factor, rel_income, index_dummy) %>%
  group_by(wealth_factor, rel_income) %>% 
  summarize(frequency = n()) %>% 
  mutate(`%` = (round(frequency/sum(frequency), 2)),
         bw = if_else(rel_income == 1, "Female breadwinner", "Not female breadwinner"),
         wealth_factor = case_when(wealth_factor == "poorest" ~ "1-poorest",
                                   wealth_factor == "poorer" ~ "2-poor",
                                   wealth_factor == "middle" ~ "3-middle",
                                   wealth_factor == "richer" ~ "4-rich",
                                   wealth_factor == "richest" ~ "5-richest")) %>% 
  #   edu_level = case_when(edu_level == "1-less than primary" ~ "Less than primary",
  #                        edu_level == "2-primary" ~ "Primary",
  #                       edu_level == "3-secondary" ~ "Secondary",
  #                      edu_level == "4-higher" ~ "Higher education")) %>% 
  ggplot(aes(x = wealth_factor,
             y = `%`,
             fill = bw, 
             label = `%`)) +
  geom_col() + 
  scale_fill_ordinal(begin = 0.2, end = 0.8) +
  theme_minimal()  +
  scale_y_continuous(limits = c(0, 0.2)) +
  labs(title = "Proportion of female breadwinners by level of household wealth \n",
       x = "",
       y = "% of female breadwinners \n",
       fill = "") +
  geom_text(  vjust = -1,
            size = 4) +
  theme(legend.position = "none")

```

```{r wealth_trad_plot, fig.height = 5, fig.width = 8}

# Wealth and traditional gender views 

data %>% 
  select(v002, wealth_factor, rel_income, index_dummy) %>%
  group_by(wealth_factor, index_dummy) %>% 
  summarize(frequency = n()) %>% 
  mutate(`%` = (round(frequency/sum(frequency), 2)),
         bw = if_else(index_dummy == 1, "Very traditional man", "Not very traditional man"),
         wealth_factor = case_when(wealth_factor == "poorest" ~ "1-poorest",
                                   wealth_factor == "poorer" ~ "2-poor",
                                   wealth_factor == "middle" ~ "3-middle",
                                   wealth_factor == "richer" ~ "4-rich",
                                   wealth_factor == "richest" ~ "5-richest")) %>% 
  #   edu_level = case_when(edu_level == "1-less than primary" ~ "Less than primary",
  #                        edu_level == "2-primary" ~ "Primary",
  #                       edu_level == "3-secondary" ~ "Secondary",
  #                      edu_level == "4-higher" ~ "Higher education")) %>% 
  ggplot(aes(x = wealth_factor,
             y = `%`,
             fill = bw, 
             label = `%`)) +
  geom_col() + 
  scale_fill_ordinal(begin = 0.2, end = 0.8, direction = -1) +
  theme_minimal()  +
  scale_y_continuous(limits = c(0, 0.2)) +
  labs(title = "Proportion of very traditional male partners by level of household wealth \n",
       x = "",
       y = "% of very traditional men \n",
       fill = "") +
  geom_text(  vjust = -1,
            size = 4) +
  theme(legend.position = "none")

```

```

```{r}

cross_edulev <- 
  prop.table(table(data$ever_control, data$factor_edulev_f), 2) %>% 
  as.data.frame() %>% arrange(Var2) %>% 
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Controlling behaviors` = round(`1`*100, 1)
        # ,`Did not experience control` = round(`0`*100, 1)
        ) %>% select(c(1,4)) %>% 
  
  left_join(prop.table(table(data$ever_emotional, data$factor_edulev_f), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Emotional violence` = round(`1`*100, 1)
         #,`Did not experience emotional violence` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2") %>% 
  
  left_join(prop.table(table(data$ever_lesssevere, data$factor_edulev_f), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Less severe violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2") %>% 
  
  left_join(prop.table(table(data$ever_severe, data$factor_edulev_f), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Severe violence` = round(`1`*100, 1)
         #,`Did not experience severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2") 
  
cross_bw <- 
  prop.table(table(data$ever_control, data$rel_income), 2) %>% 
  as.data.frame() %>% arrange(Var2) %>% 
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Controlling behaviors` = round(`1`*100, 1)
        # ,`Did not experience control` = round(`0`*100, 1)
        ) %>% select(c(1,4)) %>%
  
  left_join(prop.table(table(data$ever_emotional, data$rel_income), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Emotional violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2") %>% 
  
  left_join(prop.table(table(data$ever_lesssevere, data$rel_income), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Less severe violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2") %>% 
  
  left_join(prop.table(table(data$ever_severe, data$rel_income), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Severe violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2")

cross_wealth <- 
  prop.table(table(data$ever_control, data$wealth), 2) %>% 
  as.data.frame() %>% arrange(Var2) %>% 
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Controlling behaviors` = round(`1`*100, 1)
        # ,`Did not experience control` = round(`0`*100, 1)
        ) %>% select(c(1,4)) %>%
  
    left_join(prop.table(table(data$ever_emotional, data$wealth), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Emotional violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2") %>% 
  
  left_join(prop.table(table(data$ever_lesssevere, data$wealth), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Less severe violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2") %>% 
  
  left_join(prop.table(table(data$ever_severe, data$wealth), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Severe violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2")

cross_att <- 
  prop.table(table(data$ever_control, data$index_dummy), 2) %>% 
  as.data.frame() %>% arrange(Var2) %>% 
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Controlling behaviors` = round(`1`*100, 1)
        # ,`Did not experience control` = round(`0`*100, 1)
        ) %>% select(c(1,4)) %>%
  
   left_join(prop.table(table(data$ever_emotional, data$index_dummy), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Emotional violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2") %>% 
  
  left_join(prop.table(table(data$ever_lesssevere, data$index_dummy), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Less severe violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2") %>% 
  
  left_join(prop.table(table(data$ever_severe, data$index_dummy), 2) %>%
  as.data.frame() %>% arrange(Var2) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>% 
  mutate(`Severe violence` = round(`1`*100, 1)
         #,`Did not experience less severe v` = round(`0`*100, 1)
         ) %>% select(c(1,4)), by = "Var2")


ind_variables <- c("Less than primary",
                   "Primary",
                   "Secondary",
                   "Higher education",
                   "Woman does not earn more than man",
                   "Woman earns more than man",
                   "Not very traditional male partner (bottom 90%)",
                   "Very traditional male partner (top 10%)")

cross_table <- 
  rbind(cross_edulev,
        cross_bw,
        cross_wealth,
        cross_att) %>% 
  mutate(Subgroup = ind_variables) %>% relocate(Subgroup)  %>%  select(-Var2) %>% 
  kableExtra::kbl(caption = "Cross tabulation between dependent and independent variables* (percentage of women who have experienced violence in last 12 months)",
   #   format = "latex",
      align = "lrrrr") %>% 
  kableExtra::kable_classic_2(latex_options =c("scale_down")) %>%
  kableExtra::pack_rows("Female education level", 1,4) %>% 
  kableExtra::pack_rows("Female breadwinners", 5, 6) %>% 
  kableExtra::pack_rows("Wealth", 7, 11) %>% 
  kableExtra::pack_rows("Traditional gender attitudes", 12, 13) %>% 
  kableExtra::add_footnote(c("Wealth is not included as an independent variable in the logistic regression models but it is added here for reference"), notation = "symbol")

cross_table

```
